Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2024-06-21 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@

In the previous section, we simulated data to assess invariant power transformations and their robustness against outliers. In this section, we assess these aspects using real data.

### Top Gear dataset

@@

### Ischemic stroke dataset

@@

.plot_ischemic_stroke(
  plot_theme = plot_theme
)
```
The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
These patients underwent Computed Tomography Angiography to characterize the carotid artery blockages.
The angiography imaging was then assessed, and various characteristics related to the blood vessels and the disease are measured.
The maximum arterial wall thickness feature contains several instances with outlier values.
Figure \ref{fig:ischemic-stroke} show that non-robust transformation fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively). Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).

Datasets that where invariance has better properties in terms of lambda and resulting mean and standard deviation.
modeldata::ischemic_stroke$age
modeldata::ames$Longitudinal

>>>>

Invariant power transformations are intended to yield improved transformations to normality
in the presence of large shifts in location, scales far from unity, or both.
We previously demonstrated this property using simulations, and will now demonstrate
it using examples from real datasets.

### Age in a ...

In health-related datasets patient age is a common characteristic.

Show invariance to units (days vs years)

### Body mass of penguins
modeldata::penguins$body_mass_g

### Longitudinal coordinates in the Ames housing dataset

Datasets where invariance has better properties in terms of lambda and resulting mean and standard deviation.
modeldata::ischemic_stroke$age
modeldata::ames$Longitudinal

@@

We previously simulated data to assess invariant power transformations and their robustness against outliers. In this section, we assess these aspects using real data.

### Fuel efficiency in the Top Gear dataset

@@

### Maximum arterial wall thickness in an ischemic stroke dataset

@@

.plot_ischemic_stroke(
  plot_theme = plot_theme
)
```
The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
These patients underwent Computed Tomography Angiography to characterize the carotid artery blockages.
Angiography imaging was then assessed, and various characteristics related to the blood vessels and the disease are measured.
The maximum arterial wall thickness feature contains several instances with outlier values.
Figure \ref{fig:ischemic-stroke} show that non-robust transformation fail to produce a centrally normal distribution (empirical central normality test $p = 0.01$ and $p=0.02$ for conventional and invariant transformations, respectively). Robust transformations produce distributions that are centrally normal (empirical central normality test $p > 0.05$).
